### 2023-09-06 article/article.qmd
add initial steps for article

---
title: "SDG contributions"
format: html
---

```{r load data}
library(sparklyr)
library(ggplot2)
library(patchwork)
library(dplyr)
library(arrow)
library(here)
library(tidyr)
library(forcats)

theme_set(theme_bw())

source(here("R/helpers.R"))

Sys.setenv(SPARK_HOME = "/home/tklebel/spark-3.4.0-bin-hadoop3/")
Sys.setenv(HADOOP_HOME = "/home/hadoop/hadoop-3.3.1")
Sys.setenv(HADOOP_CONF_DIR = "/home/hadoop/hadoop-3.3.1/etc/hadoop")
Sys.setenv(YARN_HOME = "/home/hadoop/hadoop-3.3.1")
Sys.setenv(YARN_CONF_DIR = "/home/hadoop/hadoop-3.3.1/etc/hadoop")
Sys.setenv(JAVA_HOME="/usr/lib/jvm/java-1.11.0-openjdk-amd64")

config <- spark_config()
config$spark.executor.cores <- 5
config$spark.executor.instances <- 38
config$spark.executor.memory <- "20G"
config$spark.hadoop.mapreduce.fileoutputcommitter.algorithm.version <- 2

sc <- spark_connect(master = "yarn", config = config,
                    app_name = "SDG-knowledge-production")

message("Connection to Spark successful!")

message("Reading the datasets...")

paper_cols <- c("id", "rank", "doi", "type", "title_normalised", "title", 
                "unknown1", "year", "date", "publisher", paste0("unknown", 2:11),
                "journal", paste0("unknown", c(13, 14)))

mag_2020_papers <- spark_read_csv(sc, "/tklebel/mag2020/papers.txt",
                                   name = "mag2020_papers",
                                   memory = FALSE,
                                  delimiter = "\\t", header = FALSE,
                                  columns = paper_cols)

papers <- spark_read_csv(sc, "/tklebel/SDG/sdg_papers_collated.csv",
                         name = "papers")
sdg_labels <- spark_read_csv(sc, "/tklebel/SDG/sdg_labels.csv",
                             name = "sdg_labels")
# join the labels here, since we are doing everything by SDG
papers <- papers %>% 
  left_join(sdg_labels)

# remove 2020 altogether
papers <- papers %>% 
  filter(year < 2020)
```

# Growth of literature
```{r sdg_count, fig.width=7, fig.height=4}
sdg_counts <- papers %>% 
  count(SDG_label, year) %>% 
  collect()

overall_count <- mag_2020_papers %>% 
  filter(year %in% 2006:2019) %>% 
  count(year, name = "overall_n") %>% 
  collect() %>% 
  mutate(year = as.integer(year))

merged_count <- sdg_counts %>% 
  left_join(overall_count) %>% 
  mutate(prop = n / overall_n)

merged_count %>% 
  drop_na() %>% 
  ggplot(aes(as_year(year), prop, colour = fix_sdg(SDG_label))) +
  geom_line() +
  geom_point() +
  colorspace::scale_color_discrete_qualitative() +
  scale_y_log10(labels = scales::comma,
                breaks = c(10e+3, 30e+3, 1e+5, 1e+6)) +
  labs(x = NULL, y = "# of publications", colour = NULL,
       title = "Development of SDG areas over time") 
```

Normalised by the overall number of papers in MAG, there is some growth for 
these literatures.

TODO: make a better figure/table to view the growth rate

### 2023-09-06 R/helpers.R

@@ -1,0 +1,35 @@
+make_author_groups <- function(spark_author_paper_affiliations) {
+  spark_author_paper_affiliations %>%
+    group_by(paperid) %>%
+    mutate(paper_author_cat = case_when(
+      max(authorsequencenumber) == 1 ~ "single",
+      max(authorsequencenumber) == 2 ~ "double",
+      TRUE ~ "multi"
+    )) %>%
+    mutate(author_position = case_when(
+      paper_author_cat == "single" ~ "first_author",
+      paper_author_cat == "double" & authorsequencenumber == 1 ~ "first_author",
+      paper_author_cat == "double" & authorsequencenumber == 2 ~ "last_author",
+      paper_author_cat == "multi" & authorsequencenumber == 1 ~ "first_author",
+      paper_author_cat == "multi" &
+        authorsequencenumber == max(authorsequencenumber, na.rm = TRUE) ~ "last_author",
+      TRUE ~ "middle_author"
+    ))
+}
+
+# approach from https://stackoverflow.com/a/11728547/3149349
+cut_quantiles <- function(x) {
+  cut(x, breaks = quantile(x, probs = seq(0, 1, by = .2), na.rm = TRUE),
+      labels = c("p[0,20]", "p(20,40]", "p(40,60]", "p(60,80]", "p(80,100]"),
+      include.lowest = TRUE)
+}
+
+cut_quartiles <- function(x) {
+  cut(x, breaks = quantile(x, probs = seq(0, 1, by = .25), na.rm = TRUE),
+      labels = c("p[0,25]", "p(25,50]", "p(50,75]", "p(75,100]"),
+      include.lowest = TRUE)
+}
+
+fix_sdg <- function(x) fct_relevel(x, "SDG_13", after = 3)
+
+as_year <- function(num_val) lubridate::ymd(num_val, truncated = 2L)

### 2023-09-06 article/article.qmd
remove 2019 altogether and add overall growth stats
reason for removing 2019: there was a substantial drop in publications in MAG in 2019 to the year before (-13%), which means the sample is not complete there. Calculating growth rates would thus be incorrect

@@
 
-# remove 2020 altogether
+# remove 2019-2020 altogether. my MAG copy was still incomplete for 2019
 papers <- papers %>% 
-  filter(year < 2020)
+  filter(year < 2019)
+```
+
+# Sample size
+```{r}
+# total sample size
+papers %>% 
+  sdf_nrow()
+```
+
+```{r}
+# per SDG
+papers %>% 
+  count(SDG_label)
+```
+
+```{r}
+papers %>% 
+  group_by(year) %>% 
+  filter(year %in% c(2006, 2018)) %>% 
+  count(SDG_label) %>% 
+  arrange(SDG_label, year)
 ```
@@
 
-TODO: make a better figure/table to view the growth rate
+```{r growth-rates}
+growth_rate_data <- merged_count %>% 
+  arrange(SDG_label, year) %>% 
+  group_by(SDG_label) %>% 
+  mutate(prop_change = prop / lag(prop),
+         prop_change = replace_na(prop_change, 1),
+         prop_change = prop_change - 1)
+
+growth_rate_data %>% 
+  ggplot(aes(as_year(year), prop_change, colour = fix_sdg(SDG_label))) +
+  geom_line() +
+  geom_point() +
+  colorspace::scale_color_discrete_qualitative() +
+  scale_y_continuous(labels = scales::percent) +
+  labs(x = NULL, y = "Yearly growth rate relative to total papers in MAG",
+       colour = NULL,
+       title = "Development of SDG areas over time") +
+  theme(legend.position = "top")
+```
+
+```{r}
+growth_rate_data %>% 
+  filter(year %in% c(2006, 2018)) %>% 
+  summarise(overall_growth_factor = {overall_n / lag(overall_n)}[2],
+            individual_growth_factor = {n / lag(n)}[2],
+            individual_growth_rate = individual_growth_factor^(1/(2018-2006)) - 1,
+            overall_growth_rate = overall_growth_factor^(1/(2018-2006)) - 1) %>% 
+  knitr::kable()
+
+```
+
+```{r}
+spark_disconnect(sc)
+```

### 2023-09-06 article/article.qmd
add code to reproduce figure 1

@@
 library(forcats)
+library(readr)
 
@@
 
+# Overall MAG data
 paper_cols <- c("id", "rank", "doi", "type", "title_normalised", "title", 
@@
 
+# SDG specific data
 papers <- spark_read_csv(sc, "/tklebel/SDG/sdg_papers_collated.csv",
@@
   filter(year < 2019)
+
+# Affiliations
+affils <- spark_read_csv(sc,
+                         "/tklebel/SDG/affiliations_with_country_code.csv",
+                         name = "affils")
+
+author_paper_affiliations <- spark_read_csv(
+  sc,
+  "/tklebel/SDG/sdg_author_paper_affil.csv",
+  name = "author_paper_affiliations"
+)
+
+# UN country information
+un_countries <- read_csv2(here::here("data/external/UNSD — Methodology.csv"))
+
+un_countries_selection <- un_countries %>% 
+  select(country_code = `ISO-alpha3 Code`, country_name = `Country or Area`, 
+         continent = `Region Name`, sub_continent = `Sub-region Name`)
+
 ```
@@
 
+# Continents over time
+```{r}
+author_paper_affiliations_w_groups <- make_author_groups(author_paper_affiliations)
+
+papers_w_affils <- papers %>% 
+  select(paperid, SDG_label, year, citations_norm) %>% 
+  left_join(author_paper_affiliations_w_groups) %>% 
+  right_join(affils) %>% 
+  select(paperid, SDG_label, year, authorid, affiliationid, country,
+         paper_author_cat, author_position, citations_norm)
+```
+
+```{r}
+papers_per_country_fos_author_pos <- papers_w_affils %>% 
+  group_by(country, year, SDG_label, author_position) %>% 
+  summarise(n = n()) %>% 
+  collect()
+
+papers_per_country_fos_author_pos_country <- papers_per_country_fos_author_pos %>%
+  left_join(un_countries_selection, by = c("country" = "country_code")) %>% 
+  drop_na()
+
+normalised_region_share_time <- papers_per_country_fos_author_pos_country %>% 
+  filter(author_position == "first_author") %>%
+  group_by(SDG_label, year, sub_continent) %>% 
+  summarise(nn = sum(n)) %>% 
+  mutate(share = nn/sum(nn))
+```
+
+```{r sub_continents_over_time, fig.width=8.2, fig.height=4.4}
+date_scale <- scale_x_date(breaks = as_year(c(2006, 2010, 2015, 2019)),
+               date_labels = "%Y")
+base_col = "grey80"
+
+subset_countries <- filter(
+  normalised_region_share_time, 
+  sub_continent %in% c("Northern America", "Western Europe", "Northern Europe",
+                       "Eastern Asia"
+  ))
+
+normalised_region_share_time %>% 
+  ggplot(aes(as_year(year), share, group = sub_continent, colour = sub_continent)) +
+  geom_line(colour = base_col) +
+  geom_line(data = subset_countries) +
+  geom_point(data = subset_countries, size = 1.2) +
+  facet_wrap(vars(fix_sdg(SDG_label))) +
+  date_scale +
+  scale_y_continuous(labels = function(x) scales::percent(x, 1)) +
+  colorspace::scale_color_discrete_qualitative() +
+  theme(legend.position = "top") +
+  labs(x = NULL, y = "% of publications from world region within SDGs",
+       colour = NULL)
+```
+
 ```{r}

### 2023-09-06 article/article.qmd
dont restrict figure 1 to first authors.
Reason: this might overlook contributions that predominantly go to middle authors. The new approach is equivalent to full counting

@@
 normalised_region_share_time <- papers_per_country_fos_author_pos_country %>% 
-  filter(author_position == "first_author") %>%
   group_by(SDG_label, year, sub_continent) %>% 
@@
 
+This is now equivalent to full counting, since all author positions are included.
+
 ```{r}

### 2023-09-06 article/article.qmd
include 2019 again
was using an outdated file for total MAG. using the 2021 data I have gives sensible numbers of 2019

@@
 paper_cols <- c("id", "rank", "doi", "type", "title_normalised", "title", 
-                "unknown1", "year", "date", "publisher", paste0("unknown", 2:11),
-                "journal", paste0("unknown", c(13, 14)))
+                "unknown1", "year", "date", "unknown2", "publisher", 
+                paste0("unknown", 3:12),
+                "journal", paste0("unknown", c(14:16)))
 
-mag_2020_papers <- spark_read_csv(sc, "/tklebel/mag2020/papers.txt",
-                                   name = "mag2020_papers",
+mag_2021_papers <- spark_read_csv(sc, "/tklebel/mag-2021-03-15/Papers.txt",
+                                  name = "mag2021_papers",
                                   memory = FALSE,
@@
                                   columns = paper_cols)
+
+mag_2021_paper_author_affil <- spark_read_csv(
+  sc, 
+  "/tklebel/mag-2021-03-15/PaperAuthorAffiliations.txt",
+  name = "mag2021_paper_author_affil", delimiter = "\\t",
+  memory = FALSE, header = FALSE)
 
@@
 
-# remove 2019-2020 altogether. my MAG copy was still incomplete for 2019
+# remove 2020
 papers <- papers %>% 
-  filter(year < 2019)
+  filter(year < 2020)
 
@@
 
-overall_count <- mag_2020_papers %>% 
+overall_count <- mag_2021_papers %>% 
   filter(year %in% 2006:2019) %>% 
@@
 growth_rate_data %>% 
-  filter(year %in% c(2006, 2018)) %>% 
+  filter(year %in% c(2006, 2019)) %>% 
   summarise(overall_growth_factor = {overall_n / lag(overall_n)}[2],
             individual_growth_factor = {n / lag(n)}[2],
-            individual_growth_rate = individual_growth_factor^(1/(2018-2006)) - 1,
-            overall_growth_rate = overall_growth_factor^(1/(2018-2006)) - 1) %>% 
+            individual_growth_rate = individual_growth_factor^(1/(2019-2006)) - 1,
+            overall_growth_rate = overall_growth_factor^(1/(2019-2006)) - 1) %>% 
   knitr::kable()
@@
 
+Need a baseline across all of MAG. For this we need:
+
+- all papers
+- all paper author affils
+- mapping between affil and country
+
 ```{r}

### 2023-09-06 article/article.qmd
show data in output

@@
 growth_rate_data %>% 
+  knitr::kable()
+
+growth_rate_data %>% 
   filter(year %in% c(2006, 2019)) %>% 

### 2023-09-11 R/helpers.R
regather information on country of affiliations via OpenAlex

@@ -33,3 +33,49 @@ cut_quartiles <- function(x) {
 fix_sdg <- function(x) fct_relevel(x, "SDG_13", after = 3)
 
 as_year <- function(num_val) lubridate::ymd(num_val, truncated = 2L)
+
+
+
+# there is a bug in openalexR, see https://github.com/ropensci/openalexR/issues/161
+# below is the code from the package with some parts removed so it works for my
+# purposes.
+custom_import <- function (data, verbose = TRUE)
+{
+  data <- openalexR:::simple_rapply(data, `%||%`, y = NA)
+  if (!is.null(data$id)) {
+    data <- list(data)
+  }
+  n <- length(data)
+  pb <- openalexR:::oa_progress(n)
+  list_df <- vector(mode = "list", length = n)
+  institution_process <- tibble::tribble(~type, ~field, "identical",
+                                         "id", "identical", "ror", "identical", "works_api_url",
+                                         "identical", "type", "identical", "works_count", "identical",
+                                         "display_name", "identical", "country_code", "identical",
+                                         "homepage_url", "identical", "image_url", "identical",
+                                         "image_thumbnail_url", "identical", "cited_by_count",
+                                         "identical", "updated_date", "identical", "created_date",
+                                         "identical", "relevance_score", "flat", "display_name_alternatives",
+                                         "flat", "display_name_acronyms", "row_df", "geo", "rbind_df",
+                                         "counts_by_year", "rbind_df", "x_concepts", "rbind_df",
+                                         "associated_institutions", "flat", "ids")
+  for (i in seq.int(n)) {
+    if (verbose)
+      pb$tick()
+    item <- data[[i]]
+    fields <- institution_process[institution_process$field %in%
+                                    names(item), ]
+    sim_fields <- mapply(function(x, y) openalexR:::subs_na(item[[x]],
+                                                            type = y), fields$field, fields$type, SIMPLIFY = FALSE)
+
+    list_df[[i]] <- c(sim_fields)
+  }
+  col_order <- c("id", "display_name", "display_name_alternatives",
+                 "display_name_acronyms", "display_name_international",
+                 "ror", "ids", "country_code", "geo", "type", "homepage_url",
+                 "image_url", "image_thumbnail_url", "associated_institutions",
+                 "relevance_score", "works_count", "cited_by_count", "counts_by_year",
+                 "works_api_url", "x_concepts", "updated_date", "created_date")
[…]

### 2023-09-11 scripts/download-country-info.R

@@ -1,0 +1,36 @@
+# Issue: we don't have a country mapping for all countries. Presumably, because
+# we only did the mapping for affiliations in our set. This is not a huge issue,
+# since we have country information for 94.8% of affiliations, but it should be
+# 100%. We use the OpenAlex API to fill in those countries that are missing.
+# However, unfortunately we cannot look up by MAG id. we need to download all
+# affiliations and map then.
+
+library(dplyr)
+library(openalexR)
+library(readr)
+library(purrr)
+
+# Sys.setenv(openalexR.mailto = "tklebel@know-center.at")
+#
+# all_affils <- oa_request(
+#   "https://api.openalex.org/institutions?select=ids,country_code",
+#   verbose = TRUE
+# )
+#
+# write_rds(all_affils, "data/external/all_affils.rds")
+
+
+# prepare country_codes
+country_codes <- countrycode %>%
+  select(Alpha2, Alpha3) %>%
+  as_tibble()
+
+# read back and clean country information
+affils_df <- read_rds("data/external/all_affils.rds") %>%
+  custom_import() %>%
+  mutate(affiliationid = map_chr(ids, pluck, "mag", .default = "")) %>%
+  select(-ids) %>%
+  left_join(country_codes, by = join_by(country_code == Alpha2))
+
+affils_df %>%
+  write_csv("data/external/OpenAlex_affiliations.csv")

### 2023-09-11 scripts/regather-country.R

@@ -1,0 +1,68 @@
+# Issue: we don't have a country mapping for all countries. Presumably, because
+# we only did the mapping for affiliations in our set. This is not a huge issue,
+# since we have country information for 94.8% of affiliations, but it should be
+# 100%. We use the OpenAlex API to fill in those countries that are missing.
+# However, unfortunately we cannot look up by MAG id. we need to download all
+# affiliations and map then.
+
+
+library(sparklyr)
+library(dplyr)
+library(arrow)
+library(here)
+library(openalexR)
+
+Sys.setenv(SPARK_HOME = "/home/tklebel/spark-3.4.0-bin-hadoop3/")
+Sys.setenv(HADOOP_HOME = "/home/hadoop/hadoop-3.3.1")
+Sys.setenv(HADOOP_CONF_DIR = "/home/hadoop/hadoop-3.3.1/etc/hadoop")
+Sys.setenv(YARN_HOME = "/home/hadoop/hadoop-3.3.1")
+Sys.setenv(YARN_CONF_DIR = "/home/hadoop/hadoop-3.3.1/etc/hadoop")
+Sys.setenv(JAVA_HOME="/usr/lib/jvm/java-1.11.0-openjdk-amd64")
+
+config <- spark_config()
+config$spark.executor.cores <- 5
+config$spark.executor.instances <- 20
+config$spark.executor.memory <- "20G"
+config$spark.hadoop.mapreduce.fileoutputcommitter.algorithm.version <- 2
+
+sc <- spark_connect(master = "yarn", config = config,
+                    app_name = "get_country_data")
+
+message("Connection to Spark successful!")
+
+message("Reading the datasets...")
+
+# Overall MAG data
+affil_cols <- c("paperid", "authorid", "affiliationid", "authorsequencenumber",
+                "originalauthor", "originalaffiliation")
+
+mag_2021_paper_author_affil <- spark_read_csv(
+  sc,
+  "/tklebel/mag-2021-03-15/PaperAuthorAffiliations.txt",
+  name = "mag2021_paper_author_affil", delimiter = "\\t",
+  memory = FALSE, header = FALSE, columns = affil_cols)
+
+# Affiliations
+affils <- spark_read_csv(sc,
+                         "/tklebel/SDG/affiliations_with_country_code.csv",
+                         name = "affils")
+
+merged <- mag_2021_paper_author_affil %>%
+  distinct(affiliationid) %>%
+  left_join(affils) %>%
+  collect()
+
[…]

### 2023-09-11 article/article.qmd
some steps on merging country

@@
 
+affil_cols <- c("paperid", "authorid", "affiliationid", "authorsequencenumber",
+                "originalauthor", "originalaffiliation")
+
 mag_2021_paper_author_affil <- spark_read_csv(
@@
   name = "mag2021_paper_author_affil", delimiter = "\\t",
-  memory = FALSE, header = FALSE)
+  memory = FALSE, header = FALSE, columns = affil_cols)
 
@@
 ```{r}
+# check if we got countries for most affils in the full set
+merged <- mag_2021_paper_author_affil %>% 
+  distinct(affiliationid) %>% 
+  left_join(affils) %>% 
+  collect()
+
+merged_orig_data <- author_paper_affiliations %>% 
+  distinct(affiliationid) %>% 
+  left_join(affils) %>% 
+  collect()
+```
+
+```{r}
 spark_disconnect(sc)

### 2023-09-14 scripts/regather-country.R
explore content from OpenAlex and prepare for merging with main data

@@ -5,9 +5,12 @@
 # However, unfortunately we cannot look up by MAG id. we need to download all
 # affiliations and map then.
 
+# Downloading was done in "download-country-info.R". The resulting file
+# "OpenAlex_affiliations.csv" was then used here to fill in gaps.
 
 library(sparklyr)
 library(dplyr)
+library(tidyr)
 library(arrow)
 library(here)
 library(openalexR)
@@ -45,7 +48,7 @@ mag_2021_paper_author_affil <- spark_read_csv(
 # Affiliations
 affils <- spark_read_csv(sc,
                          "/tklebel/SDG/affiliations_with_country_code.csv",
-                         name = "affils")
+                         name = "affils", null_value = "NA")
 
 merged <- mag_2021_paper_author_affil %>%
   distinct(affiliationid) %>%
@@ -64,5 +67,47 @@ missing_country_affil_ids <- merged %>%
   pull(affiliationid)
 
 
-# Download all
+# read from OpenAlex data
+oa_affils <- read_csv("data/external/OpenAlex_affiliations.csv",
+                      col_types = cols(.default = col_character())) %>%
+  drop_na(affiliationid)
 
+oa_affils
+
+joined_affils <- tibble(affiliationid = missing_country_affil_ids) %>%
+  left_join(., oa_affils)
+
+joined_affils %>%
+  mutate(has_country = !is.na(country_code)) %>%
+  summarise(country_share = mean(has_country))
+# we are not at 100%, but this increases our coverage
+
+# check if our coverage has any discrepancies with the one coming from OpenAlex
+checked_countries <- merged %>%
+  left_join(oa_affils) %>%
+  filter(!is.na(country), !is.na(Alpha3)) %>%
+  mutate(same_country = country == Alpha3)
+
+checked_countries %>%
+  summarise(correct_countries = mean(same_country))
+
+checked_countries %>%
+  filter(!same_country) %>%
+  View()
+# there are some discrepancies, but mainly regarding larger corporations, where
+# it seems that different rationales for determining the country were used.
+# I stick to the countries that I had from MAG for the reason of continuity.
+# First, not to deviate too much from the methods I applied throughout, and also
+# because it is difficult to develop a rationale on which approach is better/more
+# consistent across the board (OpenAlex approach seems to be better on some
+# larger affiliations, but not uniformly)
+
+
+# export new data for further analyses
+merged %>%
+  filter(is.na(country)) %>%
[…]

### 2023-09-14 article/article.qmd
add baseline for country information

@@
 # Overall MAG data
-paper_cols <- c("id", "rank", "doi", "type", "title_normalised", "title", 
+paper_cols <- c("paperid", "rank", "doi", "type", "title_normalised", "title", 
                 "unknown1", "year", "date", "unknown2", "publisher", 
@@
   memory = FALSE, header = FALSE, columns = affil_cols)
+
+# this is data to fill in country information where we do not have it from the
+# SDG set
+missing_country_data <- spark_read_csv(
+  sc,
+  "/tklebel/SDG/additional_country_information.csv",
+  name = "missing_country_data"
+)
 
@@
   left_join(author_paper_affiliations_w_groups) %>% 
-  right_join(affils) %>% 
+  left_join(affils) %>% 
   select(paperid, SDG_label, year, authorid, affiliationid, country,
@@
 
-Need a baseline across all of MAG. For this we need:
-
-- all papers
-- all paper author affils
-- mapping between affil and country
+## Baseline calculation
 
 ```{r}
-# check if we got countries for most affils in the full set
-merged <- mag_2021_paper_author_affil %>% 
-  distinct(affiliationid) %>% 
+mag_2021_papers_w_affils <- mag_2021_papers %>% 
+  select(paperid, year) %>% 
+  left_join(mag_2021_paper_author_affil) %>% 
   left_join(affils) %>% 
+  left_join(missing_country_data) %>% 
+  select(paperid, year, authorid, affiliationid, country)
+```
+
+```{r}
+mag_2021_papers_per_country <- mag_2021_papers_w_affils %>% 
+  filter(year %in% 2006:2019) %>% 
+  group_by(country, year) %>% 
+  summarise(n = n()) %>% 
   collect()
 
-merged_orig_data <- author_paper_affiliations %>% 
-  distinct(affiliationid) %>% 
-  left_join(affils) %>% 
-  collect()
+mag_2021_papers_per_country_w_country_info <- mag_2021_papers_per_country %>%
+  left_join(un_countries_selection, by = c("country" = "country_code")) %>% 
+  drop_na()
+
+mag_2021_normalised_region_share_time <- mag_2021_papers_per_country_w_country_info %>% 
+  group_by(year, sub_continent) %>% 
+  summarise(nn = sum(n)) %>% 
+  mutate(share = nn/sum(nn))
 ```
+
+```{r baseline_sub_continents_over_time, fig.width=8.2, fig.height=4.4}
+date_scale <- scale_x_date(breaks = as_year(c(2006, 2010, 2015, 2019)),
+               date_labels = "%Y")
+base_col = "grey80"
+
+subset_countries <- filter(
+  mag_2021_normalised_region_share_time, 
+  sub_continent %in% c("Northern America", "Western Europe", "Northern Europe",
+                       "Eastern Asia"
+  ))
+
+mag_2021_normalised_region_share_time %>% 
+  ggplot(aes(as_year(year), share, group = sub_continent, colour = sub_continent)) +
+  geom_line(colour = base_col) +
+  geom_line(data = subset_countries) +
+  geom_point(data = subset_countries, size = 1.2) +
+  date_scale +
+  scale_y_continuous(labels = function(x) scales::percent(x, 1)) +
+  colorspace::scale_color_discrete_qualitative() +
+  theme(legend.position = "top") +
+  labs(x = NULL, y = "% of publications from world region across MAG",
+       colour = NULL)
+```
+
+The above is the same graph above but for all of MAG. The next step is to 
+integrate this into the existing graph. Options:
+
+- Normalise existing graph by baseline
+- Add baseline as a background (maybe with different line type)
+
+The latter seems more in line with the focus of the analysis, since the target
+is the overall distribution of knowledge production, not the relative 
+specialisation.
 

### 2023-09-15 article/article.qmd
conclude baseline calculations for world regions

@@
 
-subset_countries <- filter(
+mag_2021_subset_countries <- filter(
   mag_2021_normalised_region_share_time, 
@@
   geom_line(colour = base_col) +
-  geom_line(data = subset_countries) +
-  geom_point(data = subset_countries, size = 1.2) +
+  geom_line(data = mag_2021_subset_countries) +
+  geom_point(data = mag_2021_subset_countries, size = 1.2) +
   date_scale +
@@
 
-The above is the same graph above but for all of MAG. The next step is to 
-integrate this into the existing graph. Options:
+## Joint figure including baseline
+```{r, fig.width=8.2, fig.height=4.4}
+#| fig-cap: "Proportion of research from world regions on SDGs. The dashed line
+#|           indicates the average across all of MAG"
+#| label: fig-sub_continents_over_time_with_baseline
+normalised_region_share_time %>% 
+  ggplot(aes(as_year(year), share, group = sub_continent, colour = sub_continent)) +
+  geom_line(colour = base_col) +
+  geom_line(data = mag_2021_subset_countries, linetype = 2) +
+  geom_line(data = subset_countries) +
+  geom_point(data = subset_countries, size = 1.2) +
+  facet_wrap(vars(fix_sdg(SDG_label))) +
+  date_scale +
+  scale_y_continuous(labels = function(x) scales::percent(x, 1)) +
+  colorspace::scale_color_discrete_qualitative() +
+  theme(legend.position = "top") +
+  labs(x = NULL, y = "% of publications from world region within SDGs",
+       colour = NULL)
+```
 
-- Normalise existing graph by baseline
-- Add baseline as a background (maybe with different line type)
+```{r}
+# look at relative specialisation to identify how global south is represented in
+# SDG 2
 
-The latter seems more in line with the focus of the analysis, since the target
-is the overall distribution of knowledge production, not the relative 
-specialisation.
+continent_specialisation <- normalised_region_share_time %>% 
+  rename(nn_sdg = nn, share_sdg = share) %>% 
+  mutate(year = as.character(year)) %>% 
+  left_join(mag_2021_normalised_region_share_time) %>% 
+  mutate(specialisation = share_sdg / share)
+```
+
+```{r, fig.width=10, fig.height=10}
+#| fig-cap: "Specialisation of world regions. The specialisation factor is
+#| calulated by dividing the share of research towards a certain SDG by the 
+#| overall share of research coming from the same region. Melanesia, Polynesia
+#| and Micronesia are omitted due to low cell counts."
+continent_specialisation %>% 
+  filter(!sub_continent %in% c("Melanesia", "Polynesia", "Micronesia")) %>% 
+  ggplot(aes(as_year(year), specialisation, colour = fix_sdg(SDG_label))) +
+  geom_hline(yintercept = 1, colour = "grey30") +
+  geom_line() +
+  geom_point(size = 1.1) +
+  facet_wrap(vars(sub_continent)) +
+  colorspace::scale_color_discrete_qualitative() +
+  theme(legend.position = "top") +
+  labs(x = NULL, colour = NULL, y = "Specialisation factor")
+```
 

### 2023-09-15 article/article.qmd
add label to figure
the figure is not only about SDG 2

@@
 # look at relative specialisation to identify how global south is represented in
-# SDG 2
-
+# various SDGs
 continent_specialisation <- normalised_region_share_time %>% 
@@
 ```{r, fig.width=10, fig.height=10}
+#| label: fig-specialisation-by-world-region
 #| fig-cap: "Specialisation of world regions. The specialisation factor is

### 2023-09-15 article/article.qmd
add figures on institutional prestige (incl. baseline)

@@
 format: html
+execute:
+  keep-md: true
 ---
@@
 library(readr)
+library(stringr)
 
@@
          continent = `Region Name`, sub_continent = `Sub-region Name`)
+
+# Leiden Ranking
+leiden <- spark_read_csv(sc,
+                         "/tklebel/SDG/leiden_ranking.csv",
+                         name = "leiden")
 
@@
   theme(legend.position = "top") +
-  labs(x = NULL, colour = NULL, y = "Specialisation factor")
+  labs(x = NULL, colour = NULL, y = "Representation ratio")
+```
+
+```{r}
+continent_specialisation %>% 
+  arrange(sub_continent, SDG_label, year) %>% 
+  rmarkdown::paged_table()
+```
+
+# Institutional prestige
+```{r}
+papers_per_affiliation_per_fos <- papers %>% 
+  left_join(author_paper_affiliations) %>% 
+  left_join(affils) %>% 
+  group_by(authorid, paperid) %>% 
+  mutate(frac_value = 1/n()) %>% 
+  group_by(affiliationid, year, SDG_label) %>% 
+  summarise(n_frac_papers = sum(frac_value, na.rm = TRUE)) %>% 
+  collect()
+```
+
+```{r}
+leiden_small_local <- leiden %>%
+  filter(Field == "All sciences", Frac_counting == 1) %>%
+  select(University, Country, Period, P_top10, PP_top10, impact_P) %>% 
+  collect() %>% 
+  mutate(across(c(P_top10, PP_top10, impact_P), as.numeric),
+         last_year_of_period = str_extract(Period, "\\d{4}$"))
+
+affil_leiden_key <- read_csv2(
+  here::here("data/processed/matching_leiden.csv")
+)
+
+papers_per_affiliation_per_w_leiden <- papers_per_affiliation_per_fos %>%
+  mutate(affiliationid = as.numeric(affiliationid),
+         year = as.character(year)) %>% # needed for merging
+  left_join(affil_leiden_key) %>%
+  left_join(leiden_small_local, by = join_by(University, 
+                                             year == last_year_of_period))
+```
+
+## Inclusion into Leiden
+```{r, fig.width=9, fig.height=5}
+papers_per_affiliation_per_w_leiden %>% 
+  filter(year == "2018") %>% 
+  mutate(in_leiden = if_else(!is.na(University), "matched", "unmatched")) %>% 
+  pivot_longer(starts_with("n_"), names_to = "indicator") %>% 
+  mutate(indicator = if_else(
+    indicator == "n_frac_citations", 
+    "# of citations (fractional)",
+    "# of publications (fractional)"
+  )) %>% 
+  ggplot(aes(SDG_label, value, fill = in_leiden)) +
+  facet_wrap(vars(indicator)) + 
+  geom_boxplot() +
+  scale_y_log10(labels = scales::comma) +
+  colorspace::scale_fill_discrete_qualitative() +
+  labs(x = NULL, y = NULL, fill = "Leiden Ranking",
+       title = "Difference according to mapping to Leiden ranking") +
+  theme(legend.position = "top")
+```
+
+## Bivariate plot
+
+```{r sdg_who_ptop10_vs_sdg_production, fig.height=4, fig.width=8}
+plot_bivariate <- function(df, var, x_pos = 1200, y_pos = 11000,
+                           dv = n_frac_papers) {
+  pdata <- df %>%
+    filter(year == 2018, !is.na(impact_P))
+
+  labels <- pdata %>%
+    group_by(SDG_label) %>%
+    summarise(cor = cor({{var}}, {{dv}}, use = "pairwise.complete.obs")) %>%
+    mutate(x = x_pos, y = y_pos,
+           label = glue::glue("r = {round(cor, 2)}"))
+
+  pdata %>%
+    ggplot(aes({{var}}, {{dv}})) +
+    geom_point(alpha = .3) +
+    scale_x_log10() +
+    scale_y_log10(labels = scales::comma) +
+    facet_wrap(vars(fix_sdg(SDG_label))) +
+    geom_smooth() +
+    geom_text(data = labels, aes(x = x, y = y, label = label))
+}
+
+papers_per_affiliation_per_w_leiden %>%
+  filter(year == 2018) %>% 
+  plot_bivariate(P_top10, 800, 20000) +
+  labs(x = "# of publications of University that is in top 10% of citations 2015-2018 (fractional)",
+       y = "# of papers towards SDG in 2018 (fractional)",
+       title = "# of top 10% cited publications vs. SDG production")
+```
+
+## Over time
+```{r}
+pdata <- papers_per_affiliation_per_w_leiden %>% 
+  filter(!is.na(P_top10)) %>% 
+  group_by(Period) %>% 
+  mutate(across(c(P_top10, PP_top10, impact_P), cut_quantiles))
+
+plot_proportions <- function(df, indicator, y_var) {
+  df %>% 
+    group_by(SDG_label, year, {{indicator}}) %>% 
+    summarise(n = sum({{y_var}}, na.rm = TRUE)) %>% 
+    group_by(SDG_label, year) %>% 
+    mutate(prop = n/sum(n)) %>% 
+    ggplot(aes(as_year(year), prop, colour = {{indicator}})) +
+    geom_line() +
+    geom_point() +
+    facet_wrap(vars(fix_sdg(SDG_label))) + 
+    guides(colour = guide_legend(reverse = TRUE)) +
+    labs(x = NULL) +
+    scale_y_continuous(labels = scales::percent) +
+    colorspace::scale_color_discrete_qualitative()
+}
+```
+
+```{r sdg_who_ptop_productivity_share, fig.asp=.5, fig.width=8}
+pdata %>%
+  plot_proportions(indicator = P_top10, n_frac_papers) +
+  labs(y = "% of publications (fractional)", colour = expression(P["top 10%"]))
+```
+
+## Baseline comparison
+```{r}
+mag_2021_papers_per_affiliation_per_fos <- mag_2021_papers %>% 
+  left_join(mag_2021_paper_author_affil) %>% 
+  group_by(authorid, paperid) %>% 
+  mutate(frac_value = 1/n()) %>% 
+  group_by(affiliationid, year) %>% 
+  summarise(n_frac_papers = sum(frac_value, na.rm = TRUE)) %>% 
+  filter(!is.na(affiliationid)) %>% 
+  collect()
+
+# match with leiden data
+mag_2021_papers_per_affiliation_per_w_leiden <- mag_2021_papers_per_affiliation_per_fos %>%
+  mutate(affiliationid = as.numeric(affiliationid)) %>% # needed for merging
+  left_join(affil_leiden_key) %>%
+  left_join(leiden_small_local, by = join_by(University, 
+                                             year == last_year_of_period))
+
+# calculate baseline data to fit current plot
+baseline_data <- mag_2021_papers_per_affiliation_per_w_leiden %>% 
+  filter(!is.na(P_top10)) %>% 
+  group_by(Period) %>% 
+  mutate(across(c(P_top10, PP_top10, impact_P), cut_quantiles)) %>% 
+  group_by(year, P_top10) %>% 
+  summarise(n = sum(n_frac_papers, na.rm = TRUE)) %>% 
+  group_by(year) %>% 
+  mutate(prop = n/sum(n))
+```
+
+```{r sdg_who_ptop_productivity_share_w_baseline, fig.asp=.5, fig.width=8}
+pdata %>%
+  plot_proportions(indicator = P_top10, n_frac_papers) +
+  geom_line(data = baseline_data, linetype = 2, alpha = .7) +
+  labs(y = "% of publications (fractional)", colour = expression(P["top 10%"]))
 ```

### 2023-09-15 article/article.qmd
still need citations & should label all figures

@@
   group_by(affiliationid, year, SDG_label) %>% 
-  summarise(n_frac_papers = sum(frac_value, na.rm = TRUE)) %>% 
+  summarise(n_frac_papers = sum(frac_value, na.rm = TRUE),
+            n_frac_citations = sum(frac_value * citations_norm, na.rm = TRUE)) %>% 
   collect()
@@
 ## Inclusion into Leiden
-```{r, fig.width=9, fig.height=5}
+```{r inclusion_in_leiden, fig.width=9, fig.height=5}
 papers_per_affiliation_per_w_leiden %>% 

### 2023-09-15 article/article.qmd
add code about gender

@@
 ```
+# Gender
+Computations for gender could not be reproduced in the current repository 
+because the data are unavailable.
+The original gender-assignments were lost due to a prior server
+becoming defunct. However, since the code for all other computations was fully
+reproducible after setting up anew on a new server, I expect the results to be
+reproducible in principle, if the data were available. The code below is 
+provided for illustration, but was not run recently.
+
+```{r, eval=FALSE}
+gender_base <- papers %>% 
+  select(paperid, SDG_label, year, is_oa) %>% 
+  left_join(author_paper_affiliations_w_groups) %>% 
+  left_join(author_metadata) %>% 
+  select(paperid, SDG_label, year, author_position, authorid, 
+         paper_author_cat, gender) %>% 
+  filter(gender != "unknown", year < 2020)
+
+gender_years <- gender_base %>% 
+  group_by(SDG_label, year) %>% 
+  count(gender) %>% 
+  mutate(prop = n/sum(n)) %>% 
+  collect()
+
+gender_position <- gender_base %>% 
+  group_by(SDG_label, year, author_position) %>% 
+  count(gender) %>% 
+  mutate(prop = n/sum(n)) %>% 
+  collect()
+```
+
+```{r, eval=FALSE, fig.width=6, fig.height=3}
+# Share of female authorships by SDG.
+date_scale <- scale_x_date(breaks = as_year(c(2006, 2010, 2015, 2019)),
+               date_labels = "%Y")
+p <- gender_years %>% 
+  filter(gender == "female") %>% 
+  ggplot(aes(as_year(year), prop, colour = fix_sdg(SDG_label))) +
+  geom_line() +
+  geom_point() +
+  date_scale +
+  scale_y_continuous(labels = function(x) scales::percent(x, 1)) +
+  colorspace::scale_color_discrete_qualitative() +
+  labs(x = NULL, y = "Share of female authorships", colour = NULL) +
+  theme(legend.position = "top")
+p
+```
+
+```{r eval=FALSE, fig.width=8, fig.height=4}
+# Share of women authorships by SDG and author position.
+p <- gender_position %>% 
+  filter(gender == "female") %>% 
+  ggplot(aes(as_year(year), prop, colour = fix_sdg(SDG_label))) +
+  geom_line() +
+  geom_point() +
+  date_scale +
+  facet_wrap(vars(author_position)) +
+  scale_y_continuous(labels = function(x) scales::percent(x, 1)) +
+  colorspace::scale_color_discrete_qualitative() +
+  labs(x = NULL, y = "Share of female authorships", colour = NULL) +
+  theme(legend.position = "top")
+p
+```
+
+# Open Access
 

### 2023-09-15 article/article.qmd
add OA figures

@@
                          name = "leiden")
+
+# World Bank data
+wb_indicators <- spark_read_csv(
+  sc,
+  "/tklebel/SDG/world_bank_indicators.csv",
+  name = "wb_indicators")
+
+wb_countries <- read_csv(here::here("data/external/WDICountry.csv"))
 
@@
 ```{r}
-papers_per_affiliation_per_fos <- papers %>% 
+papers_per_affiliation_per_sdg <- papers %>% 
   left_join(author_paper_affiliations) %>% 
@@
   summarise(n_frac_papers = sum(frac_value, na.rm = TRUE),
+            n_frac_oa_papers = sum(frac_value * as.numeric(is_oa), na.rm = TRUE),
             n_frac_citations = sum(frac_value * citations_norm, na.rm = TRUE)) %>% 
@@
 
-papers_per_affiliation_per_w_leiden <- papers_per_affiliation_per_fos %>%
+papers_per_affiliation_per_SDG_w_leiden <- papers_per_affiliation_per_sdg %>%
   mutate(affiliationid = as.numeric(affiliationid),
@@
 ```{r inclusion_in_leiden, fig.width=9, fig.height=5}
-papers_per_affiliation_per_w_leiden %>% 
+papers_per_affiliation_per_SDG_w_leiden %>% 
   filter(year == "2018") %>% 
@@
 
-papers_per_affiliation_per_w_leiden %>%
+papers_per_affiliation_per_SDG_w_leiden %>%
   filter(year == 2018) %>% 
@@
 ```{r}
-pdata <- papers_per_affiliation_per_w_leiden %>% 
+pdata <- papers_per_affiliation_per_SDG_w_leiden %>% 
   filter(!is.na(P_top10)) %>% 
@@
 ```{r}
-mag_2021_papers_per_affiliation_per_fos <- mag_2021_papers %>% 
+mag_2021_papers_per_affiliation_per_sdg <- mag_2021_papers %>% 
   left_join(mag_2021_paper_author_affil) %>% 
@@
 # match with leiden data
-mag_2021_papers_per_affiliation_per_w_leiden <- mag_2021_papers_per_affiliation_per_fos %>%
+mag_2021_papers_per_affiliation_per_SDG_w_leiden <- mag_2021_papers_per_affiliation_per_sdg %>%
   mutate(affiliationid = as.numeric(affiliationid)) %>% # needed for merging
@@
 # calculate baseline data to fit current plot
-baseline_data <- mag_2021_papers_per_affiliation_per_w_leiden %>% 
+baseline_data <- mag_2021_papers_per_affiliation_per_SDG_w_leiden %>% 
   filter(!is.na(P_top10)) %>% 
@@
 # Open Access
+```{r}
+# aggregate oa_status -----
+oa_status <- papers %>%
+  filter(!is.na(is_oa)) %>%
+  select(paperid, SDG_label, year, is_oa, provider_cat)
+
+oa_per_year <- oa_status %>%
+  count(SDG_label, year, is_oa) %>%
+  collect()
+```
+
+```{r, sdg_oa_by_fos}
+oa_per_year %>%
+  group_by(SDG_label, year) %>%
+  mutate(oa_share = n/sum(n)) %>%
+  filter(is_oa) %>%
+  ggplot(aes(as_year(year), oa_share, colour = fix_sdg(SDG_label),
+             group = fix_sdg(SDG_label))) +
+  geom_point() +
+  geom_line() +
+  scale_y_continuous(labels = function(x) scales::percent(x, accuracy = 1)) +
+  colorspace::scale_color_discrete_qualitative() +
+  labs(x = NULL, y = NULL, colour = NULL) +
+  theme(legend.position = "top")
+```
+
+## By country
+```{r}
+oa_per_affiliation <- oa_status %>%
+  left_join(author_paper_affiliations) %>%
+  left_join(affils)
+
+oa_per_affiliation_selected <- oa_per_affiliation %>%
+  group_by(paperid) %>%
+  mutate(frac_count = 1 / max(authorsequencenumber, na.rm = TRUE)) %>%
+  select(paperid, authorid, is_oa, provider_cat, year, country, frac_count,
+         SDG_label)
+
+oa_per_country <- oa_per_affiliation_selected %>%
+  filter(year >= 2015 & year <= 2018) %>% 
+  group_by(country, is_oa) %>%
+  summarise(sum_frac_oa = sum(frac_count)) %>%
+  mutate(prop_oa = sum_frac_oa/sum(sum_frac_oa),
+         sum_frac_total = sum(sum_frac_oa)) %>%
+  collect()
+```
+
+```{r oa_sdg_per_country_gdp_p_cap, fig.width=9}
+wb_local <- wb_indicators %>%
+  filter(year >= 2015 & year <= 2018) %>%
+  group_by(country_name, country_code, indicator_code, indicator_name) %>% 
+  summarise(value = mean(value, na.rm = TRUE)) %>% 
+  collect()
+
+proper_countries <- wb_countries %>% 
+  filter(!is.na(`Currency Unit`)) %>% 
+  select(country_code = `Country Code`, short_name = `Short Name`,
+         region = Region, income_group = `Income Group`)
+
+oa_with_gdp_per_cap <- oa_per_country %>%
+  left_join(wb_local, by = c("country" = "country_code")) %>%
+  select(-indicator_name) %>%
+  filter(indicator_code %in% c("NY.GDP.PCAP.KD")) %>%
+  pivot_wider(names_from = indicator_code, values_from = value) %>%
+  drop_na() %>%
+  filter(is_oa)
+
+p <- oa_with_gdp_per_cap %>%
+  filter(sum_frac_total >= 50) %>%
+  left_join(proper_countries, by = c("country" = "country_code")) %>% 
+  ggplot(aes(NY.GDP.PCAP.KD, prop_oa)) +
+  geom_point(aes(size = sum_frac_total)) +
+  labs(x = "GDP per capita", y = "% of publications which are OA",
+       size = "# of publications") +
+  scale_size_continuous(trans = "sqrt", labels = scales::comma) +
+  scale_y_continuous(labels = scales::percent) +
+  scale_x_log10(breaks = c(1e+03, 2e+03, 5e+03, 1e+04, 2e+04, 5e+04, 1e+05),
+                labels = function(x) scales::comma(x, prefix = "$")) +
+  theme_bw()
+
+p1 <- p +
+  aes(colour = region)
+
+p1 +
+    geom_smooth(aes(colour = NULL), alpha = .3, show.legend = FALSE,
+                colour = "grey30") +
+  labs(colour = "World region")
+```
+### split by SDG
+```{r}
+oa_per_country_per_SDG <- oa_per_affiliation_selected %>%
+  filter(year >= 2015 & year <= 2018) %>% 
+  group_by(country, SDG_label, is_oa) %>%
+  summarise(sum_frac_oa = sum(frac_count)) %>%
+  mutate(prop_oa = sum_frac_oa/sum(sum_frac_oa),
+         sum_frac_total = sum(sum_frac_oa)) %>%
+  collect()
+
+oa_sdg_with_gdp_per_cap <- oa_per_country_per_SDG %>%
+  left_join(wb_local, by = c("country" = "country_code")) %>%
+  select(-indicator_name) %>%
+  filter(indicator_code %in% c("NY.GDP.PCAP.KD")) %>%
+  pivot_wider(names_from = indicator_code, values_from = value) %>%
+  drop_na() %>%
+  filter(is_oa)
+```
+
+```{r oa_by_sdg, fig.height=7, fig.width=9}
+oa_sdg_with_gdp_per_cap %>%
+  filter(sum_frac_total >= 50) %>%
+  left_join(proper_countries, by = c("country" = "country_code")) %>% 
+  ggplot(aes(NY.GDP.PCAP.KD, prop_oa)) +
+  geom_point(aes(size = sum_frac_total, colour = region)) +
+  geom_smooth(alpha = .3, show.legend = FALSE, colour = "grey30") +
+  labs(x = "GDP per capita", y = "% of publications which are OA",
+       colour = "World region", size = "# of publications") +
+  scale_size_continuous(trans = "sqrt", labels = scales::comma) +
+  scale_y_continuous(labels = scales::percent) +
+  scale_x_log10(breaks = c(1e+03, 2e+03, 5e+03, 1e+04, 2e+04, 5e+04, 1e+05),
+                labels = function(x) scales::comma(x, prefix = "$")) +
+  theme_bw() +
+  facet_wrap(vars(fct_relevel(SDG_label, "SDG_13", after = 3)), nrow = 2) +
+  theme(legend.position = c(.8, .2))
+```
+
+### Correlate with institutional prestige
+```{r}
+affil_oa <- papers_per_affiliation_per_SDG_w_leiden %>% 
+  mutate(oa_share = n_frac_oa_papers / n_frac_papers)
+```
+
+```{r oa_prestige_correlation, fig.width=8, fig.height=4}
+affil_oa %>% 
+  group_by(year, SDG_label) %>% 
+  filter(n_frac_papers > 50) %>% 
+  summarise(cor = cor(oa_share, P_top10, use = "pairwise.complete.obs")) %>% 
+  ggplot(aes(year, cor, colour = fix_sdg(SDG_label))) +
+  geom_line() +
+  geom_point() +
+  colorspace::scale_color_discrete_qualitative() +
+  scale_y_continuous(limits = c(0, .6)) +
+  labs(y = expression(paste("Correlation between OA shares in SDG & P"["top 10%"])),
+       x = NULL, colour = NULL) +
+  theme(legend.position = "top")
+```
 

### 2023-09-15 article/article.qmd
WS, typo & figure tweaks

@@
 ```
-### split by SDG
+### Split by SDG
 ```{r}
@@
 
-```{r oa_prestige_correlation, fig.width=8, fig.height=4}
+```{r oa_prestige_correlation, fig.width=6.5, fig.height=3.7}
 affil_oa %>% 
   group_by(year, SDG_label) %>% 
-  filter(n_frac_papers > 50) %>% 
+  filter(n_frac_papers > 50, !is.na(P_top10)) %>% 
   summarise(cor = cor(oa_share, P_top10, use = "pairwise.complete.obs")) %>% 
-  ggplot(aes(year, cor, colour = fix_sdg(SDG_label))) +
+  ggplot(aes(as_year(year), cor, colour = fix_sdg(SDG_label))) +
   geom_line() +
@@
   labs(y = expression(paste("Correlation between OA shares in SDG & P"["top 10%"])),
-       x = NULL, colour = NULL) +
-  theme(legend.position = "top")
+       x = NULL, colour = NULL) 
 ```

### 2023-11-03 R/helpers.R
add helper for income groups

@@ -32,6 +32,11 @@ cut_quartiles <- function(x) {
 
 fix_sdg <- function(x) fct_relevel(x, "SDG_13", after = 3)
 
+fix_income_group <- function(x) {
+  fct_relevel(x, "High income", "Upper middle income", "Lower middle income",
+              "Low income")
+}
+
 as_year <- function(num_val) lubridate::ymd(num_val, truncated = 2L)
 
 

### 2023-11-03 article/article.qmd
add figures on income groups

@@
 
+wb_countries_selection <- wb_countries %>% 
+  select(country_code = `Country Code`, region = Region, name = `Short Name`,
+         income_group = `Income Group`) %>% 
+  drop_na()
 ```
@@
   rmarkdown::paged_table()
+```
+
+## Income regions
+```{r}
+papers_per_country_fos_author_pos_income_region <- papers_per_country_fos_author_pos %>%
+  left_join(wb_countries_selection, by = c("country" = "country_code")) %>% 
+  drop_na()
+
+normalised_income_share_time <- papers_per_country_fos_author_pos_income_region %>% 
+  group_by(SDG_label, year, income_group) %>% 
+  summarise(nn = sum(n)) %>% 
+  mutate(share = nn/sum(nn))
+
+mag_2021_papers_per_country_w_income_info <- mag_2021_papers_per_country %>%
+  left_join(wb_countries_selection, by = c("country" = "country_code")) %>% 
+  drop_na()
+
+mag_2021_normalised_income_share_time <- mag_2021_papers_per_country_w_income_info %>% 
+  group_by(year, income_group) %>% 
+  summarise(nn = sum(n)) %>% 
+  mutate(share = nn/sum(nn))
+```
+
+```{r, fig.width=8.2, fig.height=4.4}
+#| fig-cap: "Proportion of research from income regions on SDGs. The dashed line
+#|           indicates the average across all of MAG"
+#| label: fig-income_regions_over_time_with_baseline
+normalised_income_share_time %>% 
+  ggplot(aes(as_year(year), share, group = fix_income_group(income_group), 
+             colour = fix_income_group(income_group))) +
+  geom_line() +
+  geom_line(data = mag_2021_normalised_income_share_time, linetype = 2) +
+  facet_wrap(vars(fix_sdg(SDG_label))) +
+  date_scale +
+  scale_y_continuous(labels = function(x) scales::percent(x, 1)) +
+  colorspace::scale_color_discrete_qualitative() +
+  theme(legend.position = "top") +
+  labs(x = NULL, y = "% of publications from world region within SDGs",
+       colour = NULL)
+```
+
+### Income region specialisation
+```{r}
+# look at relative specialisation to identify how global south is represented in
+# various SDGs
+income_group_specialisation <- normalised_income_share_time %>% 
+  rename(nn_sdg = nn, share_sdg = share) %>% 
+  mutate(year = as.character(year)) %>% 
+  left_join(mag_2021_normalised_income_share_time) %>% 
+  mutate(specialisation = share_sdg / share)
+```
+
+```{r, fig.width=6, fig.height=5}
+#| label: fig-specialisation-by-income-group
+#| fig-cap: "Specialisation of income groups. The specialisation factor is
+#| calulated by dividing the share of research towards a certain SDG by the 
+#| overall share of research coming from the same income groups."
+income_group_specialisation %>% 
+  ggplot(aes(as_year(year), specialisation, colour = fix_sdg(SDG_label))) +
+  geom_hline(yintercept = 1, colour = "grey30") +
+  geom_line() +
+  geom_point(size = 1.1) +
+  facet_wrap(vars(fix_income_group(income_group))) +
+  colorspace::scale_color_discrete_qualitative() +
+  theme(legend.position = "top") +
+  labs(x = NULL, colour = NULL, y = "Representation ratio")
 ```

### 2023-11-04 article/article.qmd
inspect top contributing institutions

@@
 
+# Top institutions per SDG
+```{r}
+check_top_affils <- papers %>% 
+  select(paperid, SDG_label, year, citations_norm) %>% 
+  left_join(author_paper_affiliations_w_groups) %>% 
+  left_join(affils) %>% 
+  group_by(SDG_label, displayname, rank) %>% 
+  summarise(n = n()) %>% 
+  arrange(SDG_label, desc(n)) %>% 
+  collect()
+```
+
+Interactively comparing institutions shows that there is specialisation across
+the three SDGs: certain institutions (e.g. Mayo Clinic) publish a lot in one
+SDG (here: SDG 3), but much less in others.
+
 # Institutional prestige

### 2023-11-04 scripts/01-samples-selection/02-gather-article-data.py
rename, add, and organise scripts

@@ -1,0 +1,127 @@
+import sys
+import logging
+from os import path
+import pyspark.sql.functions as f
+from pyspark.sql.functions import when
+
+
+def analyze(ss, cfg):
+    """
+    Run job
+    :param ss: SparkSession
+    :param cfg: app configuration
+    :return: None
+    """
+
+    logger = logging.getLogger(__name__)
+    logger.info('Python version: {}'.format(sys.version))
+    logger.info('Collect all information that is related to one paper')
+
+    # avoid nazis
+    spark = ss
+
+    # MAG dataset to use
+    db_name = cfg['mag_db_name']
+
+    logger.info('Reading the tables')
+
+    # read our papers
+    paper_path = path.join(cfg['hdfs']['onmerrit_dir'], "sdg_papers.csv")
+    sdg_papers = spark.read.csv(paper_path, header=True)
+
+    # read unpaywall data
+    unpaywall = spark.read \
+        .parquet("hdfs:///project/core/unpaywall/unpaywall.parquet") \
+        .select('doi', 'is_oa', 'oa_status', 'has_repository_copy',
+                'oa_locations')
+
+    funder_data = spark.read \
+        .csv("/project/core/openaire_funders/openaire_funders_clean.csv",
+             header=True)
+
+    journal_averages = spark.read \
+        .csv("/project/core/bikash_dataset/journal_averages.csv", header=True) \
+        .select("journalid", "year", "mean_citations")
+
+    conferences_df = spark \
+        .table(db_name + '.conferenceinstances')
+
+    # average citations per conference instance
+    conferences_df = conferences_df \
+        .withColumn("mean_conf_citations",
+                    conferences_df.citationcount / conferences_df.papercount) \
+        .select("conferenceinstanceid", "mean_conf_citations")
+
+
+
+    # join with unpaywall -------
+    logger.info('Joining with unpaywall')
+    # prepare new OA categorisation first
+    explodeDF = unpaywall.select(f.explode("oa_locations").alias("oa_loc"),
+                                 'doi', 'is_oa', 'oa_status', 'has_repository_copy')
+
+    flattenDF = explodeDF.selectExpr("oa_loc.is_best", "oa_loc.host_type",
+                                     'doi', 'is_oa', 'oa_status', 'has_repository_copy')
+
+    best_locations = flattenDF.distinct().filter(f.col('is_best'))
+
+    unpaywall_cat = best_locations \
+        .withColumn("provider_cat",
[…]

### 2023-11-09 article/article.qmd
compute average specialisation factor across time for income regions
fix figure width and reduce messages

@@
 
-```{r load data}
+```{r load data, message=FALSE}
 library(sparklyr)
@@
 
+```{r}
+income_group_specialisation %>% 
+  arrange(income_group, SDG_label, year) %>% 
+  rmarkdown::paged_table()
+```
+
+```{r}
+income_group_specialisation %>% 
+  group_by(income_group, SDG_label) %>% 
+  summarise(mean_specialisation = mean(specialisation))
+```
+
 # Top institutions per SDG
@@
 
-```{r oa_by_sdg, fig.height=7, fig.width=9}
+```{r oa_by_sdg, fig.height=7, fig.width=10}
 oa_sdg_with_gdp_per_cap %>%
